---
title: "MachinelearningAssignment"
output: html_document
---

##Background
Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement – a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset). 

##Data
The training data for this project are available here: https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv
The test data are available here: https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv

##https://www.google.nl/search?client=safari&rls=en&q=acknowledgement&ie=UTF-8&oe=UTF-8&gfe_rd=cr&ei=TgvXVeXmO4Pu8wfipqqwBw
The data for this project come from this source: http://groupware.les.inf.puc-rio.br/har. 

First downloading and loading the data into "r"
```{r, cache = TRUE}
setwd("~/Documents/Rstudio")
if(!file.exists("data")){
        dir.create("data")
}

fileUrl <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
download.file(fileUrl, destfile = "./data/training.csv", method="curl")
dateDownloaded <- date()
Data <- read.csv("./data/training.csv", na.strings=c('#DIV/0!', '', 'NA') ,stringsAsFactors = T)

fileUrl <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
download.file(fileUrl, destfile = "./data/testing.csv", method="curl")
dateDownloaded <- date()

FinalTest <- read.csv("./data/testing.csv")
```

Loading the required packages
```{r, cache=TRUE}
library(caret)
library(ggplot2)
require(plyr)
require(Hmisc)
library(randomForest)
```

##Data clean-up 
Since building a model on variables with no values is useless, let's delete those first. Then replace the leftover NA with the value zero. since a lot of variables contain a lot of NA's which are now replaced with zero, there mean is also zero lets delete those variables as well. Also lets subset on the new_window="no" since our 20test cases are only containing no. At last delete the first 7 columns of dataset, because this data has notting to do with measurements and classe determination.

```{r, cache=TRUE}
## delete rows with only NA values
dat <- Data[,colSums(is.na(Data))<nrow(Data)]

## replace every NA by 0
dat[is.na(dat)] <- 0

## subset on new_window 'no'; 
dat <- subset(dat, new_window == "no")

## delete columns with mean zero
dat <- dat[, colSums(dat != 0) > 0]

## deleting first 7 columns since this has nothing to do with the classe
dat1 <- dat[,-c(1:7)]
```

##Cross-validation
for the cross-validation we need to split the trainingset in a training and test set. This is done with the createDataPartition with a 60% of the data going into the trainingset.
```{r, cache=TRUE}
set.seed(1234)
inTrain <- createDataPartition(y=dat1$classe, p =0.6, list=FALSE)
training <- dat1[inTrain,]
testing <- dat1[-inTrain,]
```

Lets build our model. The used method is randomforest and the model is trained on all the left variables after the sample clean-up.
```{r, cache=TRUE}
modFit <- train(classe ~. , data =training, method="rf", prox = TRUE)
```

##performance of the model
first lets view the model and see how it is predicting the class of the testing dataset

```{r, cache=TRUE}
modelFit
pred <- predict(modFit, testing)
confusionMatrix(pred, testing$classe)
```

**The out of sample error on this model is `r 1-0,9928` (1 minus Accuracy)**

The sensitivity of the model for classe 
*A = 99.8% 
*B = 98.8% 
*C = 98.9% 
*D = 98.7% 
*E = 99.7%

The specificity of the model for classe
*A = 99.8% 
*B = 99.8% 
*C = 99.7% 
*D = 99.9% 
*E = 99.9%

##Prediction assignment
The model is now used to determine the classe of the 20 test cases.
```{r, cache=TRUE}
FinalTest[is.na(FinalTest)] <- 0
predFinal <- predict(modFit, FinalTest)
predFinal
```
Submitting the output, gave the result of 20 out of 20 correct.  
